---
title: "OSA PheWAS_Vanderbilt"
author: "Olivia J. Veatch"
date: "December 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required libraries
```{r eval=FALSE}
library(tidyverse)
library(PheWAS)
```

1) Identify BMI and Age at first use of OSA-related clinical code and summarize race, geneder, bmi in EHR-defined cases of OSA

1.1) To clean/pull BMI at first use of OSA code
``` {r eval=TRUE}
#Read Data pulled from Vanderbilt Synthetic Derivative resource, for more information see file with Notes
#Note for race variable: A = "Asian/Pacific", B = "Black", I = "Native American", M = "Mixed", N = "Other", U = "Unknown", W = "White"

EHRdefined.OSAcases.Vanderbilt <- read.csv(file = 'C:/Users/Veatcho/Desktop/OSA-PheWAS/Potential_OSA_cases_18_88/OSAGenetics_DATA_2018-11-07_1203.csv', header = TRUE, sep = ",", fill=TRUE, stringsAsFactors = FALSE)

EHRdefined.OSAcases.Vanderbilt <- EHRdefined.OSAcases.Vanderbilt %>%
  select(GRID=study_id, gender, race, dob) %>%
  mutate(dob = as.Date(dob, format = "%Y-%m-%d")) %>%
  mutate(DOB.YR = as.numeric(format(dob, '%Y')))

#Pull dates for OSA codes to match with BMI and age
#ICD9s for OSA: 327.20 (Organic sleep apnea, unspecified), 327.23 (Obstructive sleep apnea [adult, pediatric]), 327.29 (Other organic sleep apnea), 780.51 (Insomnia with sleep apnea), 780.53 (Hypersomnia with sleep apnea), 780.57 (Sleep apnea [NOS]), 
#ICD10s for OSA: G47.30 (Sleep apnea, unspecified), G47.33 (Obstructive sleep apnea [adult, pediatric]) and G47.39 (Other sleep apnea)
Case.ICD9Codes<-read.table(file = "C:/Users/Veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_icd9.txt", header=TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

Case.ICD10Codes<-read.table(file = "C:/Users/Veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_icd10.txt", header=TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

Case.ICD9Codes <- 
  Case.ICD9Codes %>%
  mutate(ICD9CodeDate=as.Date(CODE_DATE, format = "%Y-%m-%d"), ICD9Code=as.character(CODE)) %>%
  select(GRID, ICD9CodeDate, ICD9Code, Code.Desc=CODE_DESC, POS)

#Pull OSA-related codes
ICD9Codes.OSA <- 
  Case.ICD9Codes %>%
  filter(ICD9Code=='327.20' | ICD9Code=='327.23' | ICD9Code=='327.29' | ICD9Code=='780.51' | ICD9Code=='780.53' | ICD9Code=='780.57')

Case.ICD10Codes <- 
  Case.ICD10Codes %>%
  mutate(ICD10CodeDate=as.Date(CODE_DATE, format = "%Y-%m-%d"), Code.Desc=as.character(CODE_DESC)) %>%
  select(GRID, ICD10CodeDate, ICD10Code=CODE, Code.Desc, POS)

ICD10Codes.OSA <- 
  Case.ICD10Codes %>%
  filter(ICD10Code=='G47.30' | ICD10Code=='G47.33' | ICD10Code=='G47.39')

#This works for now to match BMI below b/c no one in the dataset appears to have ONLY and ICD10 code....will need to modify if no ICD9 codes are present in a record
ICD9_10Codes.OSA <- full_join(ICD9Codes.OSA, ICD10Codes.OSA, by="GRID")

#Note code below also works and produces the same as above; however, good to check specific codes and useful if there are no fuzzy matches for grep
#OSAICD9sCodeDesc <- 
#  Case.ICD9Codes %>%
#  filter(grepl("leep apnea", CODE_DESC)) %>%
#  group_by(GRID)

#To double check that EHR-based definition is correct uncomment code below
#vandyOSAicd9.10<-full_join(ICD9sCodes.OSA, ICD10sCodes.OSA, by= "GRID")

#vandyOSAicd9.10<-mutate(vandyOSAicd9.10, codesascharacters=as.character(V3.x))
#vandyOSAicd9.10<-mutate(vandyOSAicd9.10, codesascharacters.y=as.character(V3.y))

#vandyosaicd910distinct<-
#  vandyOSAicd9.10 %>%
#  group_by(V1) %>%
#  mutate(codes.diffdates.9=n_distinct(codesascharacters))
#vandyosaicd910distinct<-
#  vandyosaicd910distinct %>%
#  group_by(V1) %>%
#  mutate(codes.diffdates.10=n_distinct(codesascharacters.y))
#vandyosaicd910distinct <- 
#  mutate(vandyosaicd910distinct, codes.diffdates.9.10=(codes.diffdates.9+codes.diffdates.10))
  
#EHROSA.icd9.10s.vandy<- vandyosaicd910distinct %>%
#  filter(codes.diffdates.9.10>=2) %>%
#  distinct(V1, .keep_all = TRUE)

#Remove unneccessary data
#rm(EHROSA.icd9.10s.vandy)
#rm(vandyosaicd910distinct)
#rm(vandyOSAicd9.10)
#rm(vandyosaicd9)

#Find BMI at or close to first diagnosis code of OSA, note that Vandy BMI is wonky
ICD9_10Codes.OSA <- ICD9_10Codes.OSA %>%
  group_by(GRID) %>%
  arrange(ICD9CodeDate) %>%
  mutate(ICD9OSA.YR=as.numeric(format(ICD9CodeDate, '%Y'))) %>%
  arrange(ICD10CodeDate) %>%
  mutate(ICD10.YR=as.numeric(format(ICD10CodeDate, '%Y')))

CaseBMI<-read.table(file = "C:/Users/Veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_vitals_bmi_clean.txt", header = TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

#Evaluate data quality via summary stats
BMISummaryCases <- CaseBMI %>%
  summarise(meanBMI=mean(BMI_CLEAN), sdBMI=sd(BMI_CLEAN), minBMI=min(BMI_CLEAN), maxBMI=max(BMI_CLEAN), QuantileBMI.05=quantile(BMI_CLEAN, 0.05), QuantileBMI.95=quantile(BMI_CLEAN, 0.95))

#Prepare BMI data for joining with ICD data
CaseBMI <- CaseBMI %>%
  group_by(GRID) %>%
  separate(TEST_DATE, into = c('BMI.DATE', 'BMI.TIME'), sep = " ") %>%
  mutate(BMI.DATE = as.Date(BMI.DATE, format = "%Y-%m-%d")) %>%
  arrange(BMI.DATE) %>%
  mutate(BMI.YR=as.numeric(format(BMI.DATE, '%Y')))

#Join BMI to OSA ICD codes data
OSAICDswithBMI<-
  full_join(ICD9_10Codes.OSA, CaseBMI, by= c("GRID" = "GRID", "ICD9OSA.YR" = "BMI.YR"))

#Chose the median value of BMI for the year in which the OSA clinical code was used
#To drop (very likely) inaccurate entries for BMI uncomment the dplyr::filter step and comment dplyr::summarize median generation step
OSAICDswithBMIMedBMI.YR <- OSAICDswithBMI %>%
  select(GRID, ICD9CodeDate, ICD9Code, Code.Desc.x, ICD10CodeDate, ICD10Code, Code.Desc.y, ICD9OSA.YR, ICD10.YR, BMI.DATE, BMI_CLEAN, POSSIBLE_PREGNANT) %>%
  group_by(GRID, ICD9OSA.YR) %>%
  filter(!POSSIBLE_PREGNANT=='Y') %>%
  summarize(MedBMI.YR = median(as.numeric(BMI_CLEAN), na.rm = TRUE)) %>%
  #filter(BMI_CLEAN>10 & BMI_CLEAN<100) %>%
  distinct(GRID, .keep_all = TRUE)

#For BMI that was obvious outlier (i.e., <10 or >100), pulled other BMIs within year of code and replaced
R240914714BMI <- CaseBMI %>%
  filter(GRID=='R240914714')

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R240914714"), "MedBMI.YR"] <- NA

R203261333BMI <- CaseBMI %>%
  filter(GRID=='R203261333')

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R203261333"), "MedBMI.YR"] <- NA

R293891428BMI <- CaseBMI %>%
  filter(GRID=="R293891428")

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R293891428"), "MedBMI.YR"] <- 15.24

OSAICDswithBMIMedBMI.YR <- OSAICDswithBMIMedBMI.YR %>%
  filter(!is.na(MedBMI.YR))

EHRdefined.OSAcases.Vanderbilt <- full_join(EHRdefined.OSAcases.Vanderbilt, OSAICDswithBMIMedBMI.YR, by="GRID")

write.table(EHRdefined.OSAcases.Vanderbilt, file="C:/Users/veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/EHRdefined.OSAcases.Vanderbilt.txt")

#Remove unneccessary data
rm(OSAICDswithBMI)
rm(OSAICDswithBMIMedBMI.YR)
rm(R203261333BMI)
rm(R240914714BMI)
rm(R293891428BMI)
rm(CaseBMI)
rm(BMISummaryCases)
rm(Case.ICD10Codesdistinct)
rm(Case.ICD10Codes)
rm(Case.ICD9Codes)
rm(CaseOSA.AllICD9_10)
   
```

Optional 1.1) To calculate BMI from height/weght measurements use below code
``` {r eval=TRUE}
#Double checking BMI calculations using height and weight pulled from charts
#Case.Height<- read.table(file = "C:/Users/Veatcho/Desktop/OSA-PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_vitals_height_clean.txt", header = TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

#Case.Height <- Case.Height %>%
#  separate(TEST_DATE, into = c('HEIGHT.DATE', 'HEIGHT.TIME'), sep = " ") %>%
#  mutate(HEIGHT.DATE = as.Date(HEIGHT.DATE, format = "%Y-%m-%d")) %>%
#  mutate(HEIGHT.YEAR=as.numeric(format(HEIGHT.DATE, '%Y'))) %>%
#  group_by(GRID) %>%
#  arrange(HEIGHT.DATE)

#Case.Weight <- read.table(file = "C:/Users/Veatcho/Desktop/OSA-PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_vitals_weight_clean.txt", header = TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

#Case.Weight <- Case.Weight %>%
#  separate(TEST_DATE, into = c('WEIGHT.DATE', 'WEIGHT.TIME'), sep = " ") %>%
#  mutate(WEIGHT.DATE = as.Date(WEIGHT.DATE, format = "%Y-%m-%d")) %>%
#  mutate(WEIGHT.YEAR=as.numeric(format(WEIGHT.DATE, '%Y'))) %>%
#  group_by(GRID) %>%
#  arrange(WEIGHT.DATE)

#Case.Height.Weight <- inner_join(Case.Height, Case.Weight, by = c("GRID" = "GRID", "HEIGHT.YEAR" = "WEIGHT.YEAR"))

#Case.Height.Weight.BMI <- Case.Height.Weight %>%
#  mutate(Weight_kg=(WEIGHT_CLEAN*0.453592), Height_cm=(HEIGHT_CLEAN*2.54)) %>%
#  mutate(Case.BMIraw=(WEIGHT_CLEAN/HEIGHT_CLEAN)) %>%
#  mutate(Case.BMI_adjusted=(Weight_kg/(Height_cm/100))) %>%
#  mutate(Case.BMI_adjustedheight=(WEIGHT_CLEAN/(Height_cm/100))) %>%
#  mutate(Case.BMI_adjustedweight=(Weight_kg/HEIGHT_CLEAN))

#CaseBMI.calc.adj <- full_join(ICD9_10Codes.OSA [, c(1, 10)], Case.Height.Weight.BMI [, c(1, 5, 13)], by= c("GRID" = "GRID", "ICD9OSA.YR" = "HEIGHT.YEAR"))

#CaseBMI.calc.adj <- CaseBMI.calc.adj %>%
#  group_by(GRID) %>%
#  filter(!is.na(Case.BMI_adjusted)) %>%
#  distinct(GRID, .keep_all = TRUE)
```

1.2) To pull age at first OSA code
``` {r eval=TRUE}
#Identify age at first use of OSA-related clinical code
EHRdefined.OSAcases.Vanderbilt <- EHRdefined.OSAcases.Vanderbilt %>%
  mutate(AgeatDx=as.numeric(ICD9OSA.YR-DOB.YR)) %>%
  filter(AgeatDx >= 18 & AgeatDx <= 88)

#Attach demographic data to EHR-defined OSA case data, exclude individuals with missing BMI, and summarize race, gender, age, BMI
#CaseDemo<-read.table(file = "C:/Users/Veatcho/Desktop/OSA-PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_demo.txt", header = TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

AgeDemoCases <- EHRdefined.OSAcases.Vanderbilt %>%
  summarise(meanAgeatDx=mean(AgeatDx), sdAgeatDx=sd(AgeatDx), minAgeatDx=min(AgeatDx), maxAgeatDx=max(AgeatDx), QuantileAgeatDx.05=quantile(AgeatDx, 0.05), QuantileAgeatDx.95=quantile(AgeatDx, 0.95))
  
BMISummaryCases <- EHRdefined.OSAcases.Vanderbilt %>%
  filter(!is.na(MedBMI.YR)) %>%
  summarise(meanBMI=mean(MedBMI.YR), sdBMI=sd(MedBMI.YR), minBMI=min(MedBMI.YR), maxBMI=max(MedBMI.YR), QuantileBMI.05=quantile(MedBMI.YR, 0.05), QuantileBMI.95=quantile(MedBMI.YR, 0.95))

RaceDemoCases <- EHRdefined.OSAcases.Vanderbilt %>%
  count(race) %>%
  mutate(proportion.race=(n/6465))

GenderDemoCases <- EHRdefined.OSAcases.Vanderbilt %>%
  count(gender) %>%
  mutate(proportion.gender= (n/6465))

#Remove unneccessary data
rm(ICD9Codes.OSA)
rm(ICD10Codes.OSA)
rm(ICD9_10Codes.OSA)
rm(AgeDemoCases)
rm(BMISummaryCases)
rm(GenderDemoCases)
rm(RaceDemoCases)
```

2) Identify most recent BMI and age in years for EHR-defined controls
``` {r eval=TRUE}
EHRdefined.OSAcontrolsdemos.white.set1 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_demo.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)

EHRdefined.OSAcontrolsdemos.white.set1 <- EHRdefined.OSAcontrolsdemos.white.set1 %>%
  filter(RACE_CAUCASIAN==1) %>%
  select(GRID, GENDER, RACE=RACE_CAUCASIAN, DOB) %>%
  mutate(RACE=recode(RACE, "1"="W")) %>%
  mutate(DOB = as.Date(DOB, format = "%Y-%m-%d")) %>%
  mutate(DOB.YR = as.numeric(format(DOB, '%Y')))
```

3) Create code counts for ICD codes, set for cases to do for controls just rename labels and run same commands
``` {r eval+TRUE}
#Import ICD9 and ICD10 codes pulled from Synthetic Derivative at Vanderbilt
EHRdefined.OSAcontrolswhite.set1.ICD9 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd9.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)
EHRdefined.OSAcontrolswhite.set2.ICD9 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd9.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)
EHRdefined.OSAcontrolswhite.set2.ICD9 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd9.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)
EHRdefined.OSAcontrolswhite.set2.ICD9 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd9.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)
EHRdefined.OSAcontrolswhite.set2.ICD9 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd9.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)

EHRdefined.OSAcontrolswhite.set1.ICD10 <- read.table(file = 'C:/Users/Veatcho/Desktop/OSA PheWAS/Controls_white_18_44/Controls_white_18_44_icd10.txt', header = TRUE, sep = "\t", fill=TRUE, stringsAsFactors = FALSE)

#Export ICD10codes and convert back to ICD9
Controlswhiteset1.ICD10Codesdistinct <- EHRdefined.OSAcontrolswhite.set1.ICD10 %>%
  distinct(CODE) %>%
  select(CODE)
write.csv(Controlswhiteset1.ICD10Codesdistinct$CODE, file="C:/Users/veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/ControlIcd10codestoconvert.csv")

#Read back in ICD10 codes mapped back to 9
ICD10to9 <- read.table(file="C:/Users/veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/CaseIcd10codestoconvert Code Set Mapping Results.txt", header=TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)
Case.ICD10Codes <- left_join(Case.ICD10Codes, ICD10to9, by="ICD10Code")

#Join ICD9 and converted ICD10 codes by individual
Case.ICD10Codes <- Case.ICD10Codes %>%
  select(GRID, ICD9CodeDate=ICD10CodeDate, ICD9Code, Code.Desc, POS)

CaseOSA.AllICD9_10<-rbind(Case.ICD9Codes, Case.ICD10Codes)

CaseOSA.AllICD9_10 <- CaseOSA.AllICD9_10 %>%
  distinct(GRID, ICD9CodeDate, ICD9Code, .keep_all = TRUE)
  
CaseOSA.AllICD9_10counts<- CaseOSA.AllICD9_10 %>%
  group_by(GRID) %>%
  mutate(Code.DiffDates=n_distinct(ICD9Code)) %>%
  distinct(GRID, ICD9Code, .keep_all = TRUE) %>%
  filter(Code.DiffDates>=3)

#Find BMI at or close to first diagnosis code of OSA, note that Vandy BMI is wonky
ICD9_10Codes.OSA <- ICD9_10Codes.OSA %>%
  group_by(GRID) %>%
  arrange(ICD9CodeDate) %>%
  mutate(ICD9OSA.YR=as.numeric(format(ICD9CodeDate, '%Y'))) %>%
  arrange(ICD10CodeDate) %>%
  mutate(ICD10.YR=as.numeric(format(ICD10CodeDate, '%Y')))

CaseBMI<-read.table(file = "C:/Users/Veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/Potential_OSA_cases_18_88_vitals_bmi_clean.txt", header = TRUE, sep="\t", fill=TRUE, stringsAsFactors = FALSE)

#Evaluate data quality via summary stats
BMISummaryCases <- CaseBMI %>%
  summarise(meanBMI=mean(BMI_CLEAN), sdBMI=sd(BMI_CLEAN), minBMI=min(BMI_CLEAN), maxBMI=max(BMI_CLEAN), QuantileBMI.05=quantile(BMI_CLEAN, 0.05), QuantileBMI.95=quantile(BMI_CLEAN, 0.95))

#Prepare BMI data for joining with ICD data
CaseBMI <- CaseBMI %>%
  group_by(GRID) %>%
  separate(TEST_DATE, into = c('BMI.DATE', 'BMI.TIME'), sep = " ") %>%
  mutate(BMI.DATE = as.Date(BMI.DATE, format = "%Y-%m-%d")) %>%
  arrange(BMI.DATE) %>%
  mutate(BMI.YR=as.numeric(format(BMI.DATE, '%Y')))

#Join BMI to OSA ICD codes data
OSAICDswithBMI<-
  full_join(ICD9_10Codes.OSA, CaseBMI, by= c("GRID" = "GRID", "ICD9OSA.YR" = "BMI.YR"))

#Chose the median value of BMI for the year in which the OSA clinical code was used
#To drop (very likely) inaccurate entries for BMI uncomment the dplyr::filter step and comment dplyr::summarize median generation step
OSAICDswithBMIMedBMI.YR <- OSAICDswithBMI %>%
  select(GRID, ICD9CodeDate, ICD9Code, Code.Desc.x, ICD10CodeDate, ICD10Code, Code.Desc.y, ICD9OSA.YR, ICD10.YR, BMI.DATE, BMI_CLEAN, POSSIBLE_PREGNANT) %>%
  group_by(GRID, ICD9OSA.YR) %>%
  filter(!POSSIBLE_PREGNANT=='Y') %>%
  summarize(MedBMI.YR = median(as.numeric(BMI_CLEAN), na.rm = TRUE)) %>%
  #filter(BMI_CLEAN>10 & BMI_CLEAN<100) %>%
  distinct(GRID, .keep_all = TRUE)

#For BMI that was obvious outlier (i.e., <10 or >100), pulled other BMIs within year of code and replaced
R240914714BMI <- CaseBMI %>%
  filter(GRID=='R240914714')

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R240914714"), "MedBMI.YR"] <- NA

R203261333BMI <- CaseBMI %>%
  filter(GRID=='R203261333')

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R203261333"), "MedBMI.YR"] <- NA

R293891428BMI <- CaseBMI %>%
  filter(GRID=="R293891428")

OSAICDswithBMIMedBMI.YR[which(OSAICDswithBMIMedBMI.YR$GRID=="R293891428"), "MedBMI.YR"] <- 15.24

OSAICDswithBMIMedBMI.YR <- OSAICDswithBMIMedBMI.YR %>%
  filter(!is.na(MedBMI.YR))

EHRdefined.OSAcases.Vanderbilt <- full_join(EHRdefined.OSAcases.Vanderbilt, OSAICDswithBMIMedBMI.YR, by="GRID")

write.table(EHRdefined.OSAcases.Vanderbilt, file="C:/Users/veatcho/Desktop/OSA PheWAS/Potential_OSA_cases_18_88/EHRdefined.OSAcases.Vanderbilt.txt")

#Remove unneccessary data
rm(OSAICDswithBMI)
rm(OSAICDswithBMIMedBMI.YR)
rm(R203261333BMI)
rm(R240914714BMI)
rm(R293891428BMI)
rm(CaseBMI)
rm(BMISummaryCases)
```

#Setting Labels

label(data$study_id)="Study ID"
label(data$gender)="Gender"
label(data$race)="Race"
label(data$ethnicity)="Ethnicity"
label(data$dob)="Date of Birth"
label(data$deceased)="Deceased"
label(data$demographics_complete)="Complete?"
#Setting Units


#Setting Factors(will create new variable for factors)
data$gender.factor = factor(data$gender,levels=c("M","F","U"))
data$race.factor = factor(data$race,levels=c("B","A","W","H","I","U","M","N"))
data$ethnicity.factor = factor(data$ethnicity,levels=c("HL","NH","UN"))
data$deceased.factor = factor(data$deceased,levels=c("Y"))
data$demographics_complete.factor = factor(data$demographics_complete,levels=c("0","1","2"))

levels(data$gender.factor)=c("Male","Female","Unknown")
levels(data$race.factor)=c("African American","Asian/Pacific","Caucasian","Hispanic","Native American","Unknown","Multi race","Other")
levels(data$ethnicity.factor)=c("Hispanic/Latino","Not Hispanic/Latino","Unknown")
levels(data$deceased.factor)=c("Yes")
levels(data$demographics_complete.factor)=c("Incomplete","Unverified","Complete")
